Sentiment Analysis of Twitter in Relation to Fossil Fuel Stock Prices

### 2023-04-27 manuscript/manuscript.tex
Final Version Before Review

The Chevron Corporation is an America-based energy company thats has operations in over 180 countries where it has vertically integrated all operations in its supply chain such as exploration, production, refinement, chemical manufacturing, transportation, sales, marketing, and power generation. A lot of revenue comes from these streams but they also receive funding from government subsidies and stock revenue. As of when this paper was written, the Chevron Corporation has the 11th highest market capitalization with 361.57 billion dollars in traded stocks.

@@

A popular method of analyzing sequence data for most natural language processing is through the use of transformer models  \citep{vaswani2017attention}. This model sought to improve the problems existing with the Recurrent Neural Network(RNN) model which used to be the preferred model for language processing tasks. Transformers have shown a lot of promise, outperforming its competitors in speed and accuracy for natural language processing tasks. Due to its novelty and its promising qualities, many are applying Transformers for various tasks, yet nobody has this used this approach to find out whether it is possible to find out the correlation between Twitter Sentiment and Stock prices of prominent fossil fuel corporations. In this paper, we specifically will focus on the Chevron Corporation (though in future iterations, we might extend this to multiple corporations in this field).
%\paragraph{My Contribution}

@@

This following is a roadmap for this study: Section \ref{sec:intro}: Introducing Concepts and Providing Background, Section \ref{sec:datadesc}: Describing the Data , Section \ref{sec:methods}: Explaining the Statistical Methods and the Machine Learning Concepts, Section \ref{sec:results}: Explaining the Results of our Paper, and Section \ref{sec:discussion}: Conducting Discussion of this Study.

@@

For the period of 2021-2022,  CVX stock data and twitter tweets were collected for analysis.
Historic Stock data for the Chevron Corporation was collected using Yahoo Finance's Historical Stock Data tool. Figure ~\ref{fig:stockdata} shows the fluctuations of the Stock Highs, the Stock Lows and the Closing Prices over the observed period that was generated using the matplotlib package for python \citep{Hunter_2007}. 
While the Highs and Lows might be informative in analysis, in order to concentrate the focus of this paper in a manner more similar to other financial data analyses, we will only experiment the closing Price and the Returns which can be taken by taking the log difference of the closing Prices, refer to Figure~{fig:returns}.
%%\jy{the figure x-labels are cutoff}
%%% fixed this
\begin{figure}[tbp]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig1.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022.}\label{fig:stockdata}
  \end{center}
\end{figure}

@@

\jy{make sure every figure/table is referred to in the text}

@@

We also used the data processing library Pandas in order to generate summary statistics as shown in Table~\ref{tab:sumstat}
The range of the closing values for the observed period was 84.70 to 188.05; the lowest and highest values were observed on January 4th, 2021 November 15th, 2022. There are 502 days that the stock data was available for since Chevron is a part of the New York Stock Exchange which is closed all day on Saturdays, Sundays and nine holidays a year. This was something that was taken into consideration when working with twitter data.

@@

For the purpose of this paper, 10 tweets were scraped with the topic query ``CVX" from January 1st 2021 to December 31st 2022.  7290 tweets were generated, saved as a csv and merged with the CVX stock data on the Date axis after matching datatypes. This also omitted tweets which were created when the stock market was closed in order to simplify the analysis. Using this we were able to perform sentiment analysis and calculate correlation.

@@

\subsection{Sentiment Analysis}
\paragraph{The Transformer Model}
As mentioned in the first section, Transformers models have shown a lot of promise for a variety of natural language processing tasks, outperforming its competitors in speed and accuracy. The BERT model seems to be particularly well suited for tasks of sentiment analysis. Transformers work based on the concept of ``attention" which was introduced in ``Attention is all you need" \citep{vaswani2017attention} . Attention allows Transformers to learn the relationships between the words in a sentence or sequence of words. To do this, the Transformer splits the input sequence into tokens (usually individual words or sub-words) and embeds them into a high-dimensional vector space. It then applies multiple layers of self-attention, where each token in the sequence is given a weight based on its relevance to the other tokens. The self-attention mechanism allows the model to focus on the most important parts of the input sequence and learn dependencies between the words. After the self-attention layers, the Transformer applies feedforward neural networks to further refine the representations of the tokens. Transformers create differential weights signaling which words in a sentence are the most critical to further process.

\begin{figure}[tbp]
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
%Rounded Rect [id:dp4434986972980839] 
\draw   (100,43.8) .. controls (100,32.76) and (108.95,23.8) .. (120,23.8) -- (261,23.8) .. controls (272.05,23.8) and (281,32.76) .. (281,43.8) -- (281,103.8) .. controls (281,114.85) and (272.05,123.8) .. (261,123.8) -- (120,123.8) .. controls (108.95,123.8) and (100,114.85) .. (100,103.8) -- cycle ;
%Rounded Rect [id:dp8739677930569794] 
\draw   (359,68.2) .. controls (359,52.52) and (371.72,39.8) .. (387.4,39.8) -- (472.6,39.8) .. controls (488.28,39.8) and (501,52.52) .. (501,68.2) -- (501,168.4) .. controls (501,184.09) and (488.28,196.8) .. (472.6,196.8) -- (387.4,196.8) .. controls (371.72,196.8) and (359,184.09) .. (359,168.4) -- cycle ;
%Curve Lines [id:da9359490056283244] 
\draw    (282,73) .. controls (362,56.8) and (275,143.8) .. (359,124.8) ;
%Straight Lines [id:da03881732608220867] 
\draw    (78,72.8) -- (101,73) ;
%Straight Lines [id:da6091063106538472] 
\draw    (500,120.8) -- (523,121) ;
%Straight Lines [id:da19347919375178857] 
\draw    (431,215.8) -- (431,199.8) ;
% Text Node
\draw (161,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {Encoder};
% Text Node
\draw (398,103) node [anchor=north west][inner sep=0.75pt]   [align=left] {Decoder};
% Text Node
\draw (42,63) node [anchor=north west][inner sep=0.75pt]   [align=left] {Input};
% Text Node
\draw (524,112) node [anchor=north west][inner sep=0.75pt]   [align=left] {Output};
% Text Node
\draw (408,219) node [anchor=north west][inner sep=0.75pt]   [align=left] { Output Probabilites};
\end{tikzpicture}
\caption{In this figure we can observe a very simplified version of the Transformer model with its two major components, the encoder and the decoder. \citep{devlin2018bert}.}
\label{fig:transformer}
\end{figure}
The Transformer architecture consists of two main components: an encoder and a decoder.The encoder takes in the input sequence and generates a sequence of hidden representations, which are then passed to the decoder. The decoder then uses the encoder's output to generate the final output sequence. Figure~\ref{fig:transformer} is a simple layout of this architecture

@@

>>>>

The Chevron Corporation is an energy company based in the United States thats has operations in over 180 countries where it has vertically integrated all operations in its supply chain such as exploration, production, refinement, chemical manufacturing, transportation, sales, marketing, and power generation. A lot of revenue comes from these streams but the corporation also receives funding from government subsidies and stock revenue. As of when this paper was written, the Chevron Corporation has the 11th highest market capitalization with 361.57 billion dollars in traded stocks.

@@

A popular method of analyzing sequence data for most natural language processing is through the use of transformer models  \citep{vaswani2017attention}. This model sought to improve the problems existing with the Recurrent Neural Network(RNN) model which used to be the preferred model for language processing tasks. Transformers have shown a lot of promise, outperforming its competitors in speed and accuracy for natural language processing tasks. Due to its novelty and its promising qualities, many are applying Transformers for various tasks, yet nobody has this used this approach to find out whether it is possible to find out the correlation between Twitter Sentiment and Stock prices of prominent fossil fuel corporations. 
%\paragraph{My Contribution}

@@

This following is a roadmap for this study: Section \ref{sec:intro}: Introducing Concepts and Providing Background, Section \ref{sec:datadesc}: Describing the Data, Section \ref{sec:methods}: Explaining the Statistical Methods and the Machine Learning Concepts, Section \ref{sec:results}: Explaining the Results of our Paper, and Section \ref{sec:discussion}: Conducting Discussion of this Study.

@@

For the period of 2021-2022,  CVX stock data and twitter tweets were collected for analysis.
Historic Stock data for the Chevron Corporation was collected using Yahoo Finance's Historical Stock Data tool. Figure ~\ref{fig:stockdata} shows the fluctuations of the Stock Highs, the Stock Lows and the Closing Prices over the observed period that was generated using the matplotlib package for python \citep{Hunter_2007}. 
While the Highs and Lows might be informative in analysis, in order to concentrate the focus of this paper in a manner more similar to other financial data analyses, we will only experiment the closing Price and the Returns which can be taken by taking the log difference of the closing Prices, refer to Figure~\ref{fig:returns}.
%%\jy{the figure x-labels are cutoff}
%%% fixed this
\begin{figure}[tbp]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig1.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022.}\label{fig:stockdata}
  \end{center}
\end{figure}

@@

%%%\jy{make sure every figure/table is referred to in the text}

@@

We also used the data processing library Pandas in order to generate summary statistics as shown in Table~\ref{tab:sumstat}.
The range of the closing values for the observed period was 84.70 to 188.05; the lowest and highest values were observed on January 4th, 2021 November 15th, 2022. There are 502 days that the stock data was available for since Chevron is a part of the New York Stock Exchange which is closed all day on Saturdays, Sundays and nine holidays a year. This was something that was taken into consideration when working with twitter data.

@@

For the purpose of this paper, 10 tweets were scraped daily with the topic query ``CVX" from January 1st 2021 to December 31st 2022.  7290 tweets were generated, saved as a csv and merged with the CVX stock data on the Date axis after matching datatypes. This also omitted tweets which were created when the stock market was closed in order to simplify the analysis. Using this we were able to perform sentiment analysis and calculate correlation.

@@

\subsection{Sentiment Analysis}
\paragraph{The Transformer Model}
As mentioned in the first section, Transformers models have shown a lot of promise for a variety of natural language processing tasks, outperforming its competitors in speed and accuracy. The BERT model seems to be particularly well suited for tasks of sentiment analysis. Transformers work based on the concept of ``attention" which was introduced in ``Attention is all you need" \citep{vaswani2017attention}. Attention allows Transformers to learn the relationships between the words in a sentence or sequence of words. To do this, the Transformer splits the input sequence into tokens (usually individual words or sub-words) and embeds them into a high-dimensional vector space. It then applies multiple layers of self-attention, where each token in the sequence is given a weight based on its relevance to the other tokens. The self-attention mechanism allows the model to focus on the most important parts of the input sequence and learn dependencies between the words. After the self-attention layers, the Transformer applies feedforward neural networks to further refine the representations of the tokens. Transformers create differential weights signaling which words in a sentence are the most critical to further process.

\begin{figure}[tbp]
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
%Rounded Rect [id:dp4434986972980839] 
\draw   (100,43.8) .. controls (100,32.76) and (108.95,23.8) .. (120,23.8) -- (261,23.8) .. controls (272.05,23.8) and (281,32.76) .. (281,43.8) -- (281,103.8) .. controls (281,114.85) and (272.05,123.8) .. (261,123.8) -- (120,123.8) .. controls (108.95,123.8) and (100,114.85) .. (100,103.8) -- cycle ;
%Rounded Rect [id:dp8739677930569794] 
\draw   (359,68.2) .. controls (359,52.52) and (371.72,39.8) .. (387.4,39.8) -- (472.6,39.8) .. controls (488.28,39.8) and (501,52.52) .. (501,68.2) -- (501,168.4) .. controls (501,184.09) and (488.28,196.8) .. (472.6,196.8) -- (387.4,196.8) .. controls (371.72,196.8) and (359,184.09) .. (359,168.4) -- cycle ;
%Curve Lines [id:da9359490056283244] 
\draw    (282,73) .. controls (362,56.8) and (275,143.8) .. (359,124.8) ;
%Straight Lines [id:da03881732608220867] 
\draw    (78,72.8) -- (101,73) ;
%Straight Lines [id:da6091063106538472] 
\draw    (500,120.8) -- (523,121) ;
%Straight Lines [id:da19347919375178857] 
\draw    (431,215.8) -- (431,199.8) ;
% Text Node
\draw (161,65) node [anchor=north west][inner sep=0.75pt]   [align=left] {Encoder};
% Text Node
\draw (398,103) node [anchor=north west][inner sep=0.75pt]   [align=left] {Decoder};
% Text Node
\draw (42,63) node [anchor=north west][inner sep=0.75pt]   [align=left] {Input};
% Text Node
\draw (524,112) node [anchor=north west][inner sep=0.75pt]   [align=left] {Output};
% Text Node
\draw (408,219) node [anchor=north west][inner sep=0.75pt]   [align=left] { Output Probabilites};
\end{tikzpicture}
\caption{In this figure we can observe a very simplified version of the Transformer model with its two major components, the encoder and the decoder. \citep{devlin2018bert}.}
\label{fig:transformer}
\end{figure}
The Transformer architecture consists of two main components: an encoder and a decoder.The encoder takes in the input sequence and generates a sequence of hidden representations, which are then passed to the decoder. The decoder then uses the encoder's output to generate the final output sequence. Figure~\ref{fig:transformer} is a simple representation of this architecture that takes into account its key components.

@@

Similarly, z-tests were conducted for weighted sentiment as the exogenous variable and the null hypothesis was still rejected with p values of 0.476 and 1 for the ARMA and GARCH portions respectively.
